---
title: "10 Fallstudie: Churn Prediction"
author: "Martin Hanewald"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
set.seed(1337)
```


## Packages

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(lubridate)
library(caret)
```

## Überblick


## Dataset

Zwei Datensätze:

- User-Stammdaten
- Aktivitäten pro User

```{r message=FALSE}
activity_raw <-  read_csv('data/activity_info.csv')
user_raw <- read_csv('data/user_info.csv')
```

## Pre-Processing

```{r}
CHURN_PERIOD <- 21
CHURN_THRESHOLD <- 0
```

### Cleanup

```{r}
activity <- activity_raw %>% 
    mutate(Timestamp = mdy_hm(Timestamp) %>% as_date() + years(17)) %>%
    mutate(ItemId = str_pad(ItemId, 12, pad='0')) %>% 
    mutate_at(vars(TransactionId, UserId, ItemId), as.factor) %>% 
    select(-Location, -ProductCategory, -X1)

user <- user_raw %>% 
    select(-Gender, -UserType) %>% 
    mutate_all(as.factor)
```

### Create the `churn` label

```{r fig.width=7}
last_date <- max(activity$Timestamp)

# Show Plot with churn period
activity %>% 
    ggplot(aes(Timestamp)) + geom_bar() + 
    annotate(geom='rect', 
              xmin=last_date-CHURN_PERIOD, 
              ymin = 0, ymax=7000, xmax=last_date,
              fill='red', alpha=.2)
```


```{r}
# Approach one: For-loop, save results in list
churn <- list()

for(u in unique(activity$UserId)){
    churn[[u]] <- activity %>%
        filter(UserId == u) %>% 
        filter(Timestamp > last_date - CHURN_PERIOD) %>% 
        nrow() <= CHURN_THRESHOLD
}

churn_label <- churn %>% as_tibble() %>% 
    gather(UserId, churn)

# profiler: 25,56 sec
```

```{r}
# Approach two: Group_by and summarise
churn_label <- activity %>% 
    mutate(in_period = Timestamp > last_date - CHURN_PERIOD) %>% 
    select(UserId, TransactionId, in_period) %>% unique() %>% 
    group_by(UserId) %>% 
    summarise(churn = sum(in_period) <= CHURN_THRESHOLD) %>% 
    mutate_at(vars(churn), as.factor)

# profiler: 3,6 sec

churn_label$churn %>% table()

```

### Feature Engineering

- Total_Quantity
- Total_Value

- StDev_Quantity
- StDev_Value

- AvgTimeDelta
- Recency

- Count_Unique_TransactionId
- Count_Unique_ItemId

- Mean_Quantity_per_Unique_TransactionId
- Mean_Quantity_per_Unique_ItemId

- Mean_Value_per_Unique_TransactionId
- Mean_Value_per_Unique_ItemId


```{r}
activity_before <- activity %>% filter(Timestamp < last_date - CHURN_PERIOD)

activity_measures <- 
    activity_before %>% 
        group_by(UserId) %>% 
        arrange(Timestamp, .by_group=TRUE) %>% 
        summarise(Total_Quantity = sum(Quantity),
                  Total_Value = sum(Value),
                  StDev_Quantity = sd(Quantity),
                  StDev_Value = sd(Value),
                  AvgTimeDelta = mean(diff(Timestamp)),
                  Recency = last_date - CHURN_PERIOD - max(Timestamp)) %>% 
    full_join(
        activity_before %>% 
            group_by(UserId, ItemId) %>% 
            summarise(Quantity_sum = sum(Quantity),
                      Value_sum = sum(Value)) %>% 
            summarise(Mean_Quantity_per_Unique_ItemId = mean(Quantity_sum),
                      Mean_Value_per_Unique_ItemId = mean(Value_sum),
                      Count_Unique_ItemId = n_distinct(ItemId))) %>% 
    full_join(
        activity_before %>% 
            group_by(UserId, TransactionId) %>% 
            summarise(Quantity_sum = sum(Quantity),
                      Value_sum = sum(Value)) %>% 
            summarise(Mean_Quantity_per_Unique_TransactionId = mean(Quantity_sum),
                      Mean_Value_per_Unique_TransactionId = mean(Value_sum),
                      Count_Unique_TransactionId = n_distinct(TransactionId)))

```

### Bind final model dataframe

```{r}
modeldat <- user %>% 
    inner_join(churn_label) %>% 
    inner_join(activity_measures) %>%
    drop_na() # 724 lines dropped where only one transaction per user

# check for na's
modeldat %>% 
    summarise_all(function(x) is.na(x) %>% sum()) %>% 
    gather(var, na) %>% filter(na>0)
```

## Modelling

### Splitting

```{r}
inTrain <- createDataPartition(modeldat$churn, p = .8, list=F)
training <- modeldat[inTrain,]
testing <- modeldat[-inTrain,]
```

### Training

```{r, cache=T}
trControl <- trainControl(method='cv')

fit <- list()

fit$glm <- train(churn ~ . -UserId,
                method = 'glm',
                family = 'binomial',
                trControl = trControl,
                preProcess = c('center', 'scale'),
                data = training)

fit$logitboost <- train(churn ~ . -UserId,
                method = 'LogitBoost',
                trControl = trControl,
                preProcess = c('center', 'scale'),
                data = training)

fit$xgbtree <- train(churn ~ . -UserId,
                method = 'xgbTree',
                trControl = trControl,
                preProcess = c('center', 'scale'),
                data = training)

```

### Evaluation
 
```{r fig.width=7}

resamples(fit) %>% bwplot()

varImp(fit$glm)
#varImp(fit$logitboost)
varImp(fit$xgbtree)

fit$glm %>% predict(testing) %>% confusionMatrix(testing$churn)
fit$logitboost %>% predict(testing) %>% confusionMatrix(testing$churn)
fit$xgbtree %>% predict(testing) %>% confusionMatrix(testing$churn)

```
 


```{css, echo=FALSE}
img{
     border:0;
     background-color: rgba(255, 255, 255, 0);
   }
   
@media print {
  div.divFooterLeft {
    position: fixed;
    bottom: 0;
    right: 0;

  }

div.divFooterRight {
    position: fixed;
    top: 0;
    right: 0;
    border:0;
  }    
}
```

<div class="divFooterLeft"><img src='../98_assets/ca_logo.svg' width="234px"></div>
<div class="divFooterRight"><img src='../98_assets/c_by_qunis.svg' width="100px"></div>

